\Large{Adversarial Variational Optimization of Non-Differentiable Simulators}

### 2017-06-20 paper.tex
save

Complex computer simulators are increasingly used across fields of science to
describe generative models tying parameters of an underlying theory to
experimental observations. However, inference in this likelihood-free setup is often
intractable, as simulators rarely provide a way to evaluate the likehood
function for a given observation. In this note, we develop a likelihood-free
inference algorithm for fitting a forward non-differentiable generative model to
observed data. We adapt the adversarial training procedure of generative
adversarial networks by replacing the implicit generative network with a
domain-based scientific simulator, and solve the resulting non-differentiable
minimax problem with variational optimization. Effectively, the adversarial
variational optimization procedure results in learning an arbitrarily tight
proposal distribution over simulator parameters, such that the respective
hierarchical model matches the observed data.
\glnote{Mention experimental results.}
\glnote{Add 'so what?' conclusion.}

@@

In many fields of science, computer simulators are used to describe complex
processes that tie parameters of an underlying theory to high dimensional
observations. In most cases, these implicit generative
models~\citep{2016arXiv161003483M} are specified as imperative implementations
of forward stochastic processes that generate data. Because it is usually
computationally intractable, simulators often do not provide a way to directly
evaluate the likelihood function for a given observation,  which motivates a
class of algorithms for likelihood-free inference.

\glnote{State contributions clearly.}

@@

As thoroughly explored in \citep{2017arXiv170104862A}, GANs remain remarkably
difficult to train because of vanishing gradients as $d$ saturates, or because of
unreliable updates when the training procedure is relaxed. As a remedy,
Wasserstein GANs~\citep{2017arXiv170107875A} reformulate the adversarial
setup in order to minimize the Wasserstein-1 distance $W(p_r, p_\theta)$ by
replacing the adversarial classifier with a 1-Lipschitz adversarial critic
$d(\cdot; \phi) : \mathbb{R}^d \to \mathbb{R}$. Under the WGAN-GP formulation of \cite{2017arXiv170400028G}
for stabilizing the optimization procedure,
training $d$ and $g$ results in alternating gradient updates on $\phi$ and $\theta$ in order to respectively minimize
\begin{align}
    {\cal L}_d =\,& \mathbb{E}_{\tilde{\mathbf{x}} \sim p_\theta} [d(\tilde{\mathbf{x}};\phi)] - \mathbb{E}_{\mathbf{x} \sim p_r} [d(\mathbf{x};\phi)]  \nonumber \\
                  & + \lambda \mathbb{E}_{\hat{\mathbf{x}} \sim p_{\hat{\mathbf{x}}}} [(|| \nabla_{\hat{\mathbf{x}}} d({\hat{\mathbf{x}}};\phi) ||_2 - 1)^2] \\
    {\cal L}_g =\,& -\mathbb{E}_{\tilde{\mathbf{x}} \sim p_\theta} [d(\tilde{\mathbf{x}};\phi)]
\end{align}
where ${\hat{\mathbf{x}}} := \epsilon \mathbf{x} + (1-\epsilon)\tilde{\mathbf{x}}$, for $\epsilon \sim U[0,1]$, $\mathbf{x} \sim p_r$ and $\tilde{\mathbf{x}} \sim p_\theta$.

@@

@@

\glnote{Implicit generative models.}
\glnote{classifier ABC.}
\glnote{carl~\citep{cranmer2015approximating}.}
\glnote{CMA-ES.}

>>>>

Complex computer simulators are increasingly used across fields of science to
describe generative models tying parameters of an underlying theory to
experimental observations. Inference in this setup is often
difficult, as simulators rarely provide a way to evaluate the likehood
function for a given observation. In this note, we develop a likelihood-free
inference algorithm for fitting a forward non-differentiable generative model to
observed data. We adapt the adversarial training procedure of generative
adversarial networks by replacing the implicit generative network with a
domain-based scientific simulator, and solve the resulting non-differentiable
minimax problem with variational optimization. Effectively, the adversarial
variational optimization procedure results in learning an arbitrarily tight
proposal distribution over simulator parameters, such that the respective
hierarchical model matches the observed data.
\glnote{Mention experimental results.}
\glnote{Add 'so what?' conclusion.}

@@

In many fields of science such as particle physics, climatology or population
genetics, computer simulators are used to describe complex processes that tie
parameters of an underlying theory to high dimensional observations. In most
cases, these implicit generative models~\citep{2016arXiv161003483M} are
specified as imperative implementations of forward stochastic processes that
generate data. Because it is usually computationally intractable, simulators
often do not provide a way to directly evaluate the likelihood function for a
given observation, thereby making inference difficult.

In this note, we develop a likelihood-free inference algorithm for fitting a
forward non-differentiable generative model to observed data. We propose to
adapt the adversarial training procedure of generative adversarial
networks~\cite{goodfellow2014generative} by replacing the implicit generative
network with a domain-based scientific simulator, and solve the resulting
non-differentiable minimax problem with variational
optimization~\citep{2012arXiv1212.4507S}. The adversarial variational
optimization procedure results in learning a proposal distribution over
simulator parameters, hence producing an abitrarily tight family of models whose
joint collection of generated samples matches the observed data.

@@

As thoroughly explored in \citep{2017arXiv170104862A}, GANs remain remarkably
difficult to train because of vanishing gradients as $d$ saturates, or because of
unreliable updates when the training procedure is relaxed. As a remedy,
Wasserstein GANs~\citep{2017arXiv170107875A} reformulate the adversarial
setup in order to minimize the Wasserstein-1 distance $W(p_r, p_\theta)$ by
replacing the adversarial classifier with a 1-Lipschitz adversarial critic
$d(\cdot; \phi) : \mathbb{R}^d \to \mathbb{R}$. Under the WGAN-GP formulation of \cite{2017arXiv170400028G}
for stabilizing the optimization procedure,
training $d$ and $g$ results in alternating gradient updates on $\phi$ and $\theta$ in order to respectively minimize
\begin{align}
    {\cal L}_d =\,& \mathbb{E}_{\tilde{\mathbf{x}} \sim p_\theta} [d(\tilde{\mathbf{x}};\phi)] - \mathbb{E}_{\mathbf{x} \sim p_r} [d(\mathbf{x};\phi)]  \nonumber \\
                  & + \lambda \mathbb{E}_{\hat{\mathbf{x}} \sim p_{\hat{\mathbf{x}}}} [(|| \nabla_{\hat{\mathbf{x}}} d({\hat{\mathbf{x}}};\phi) ||_2 - 1)^2] \\
    {\cal L}_g =\,& -\mathbb{E}_{\tilde{\mathbf{x}} \sim p_\theta} [d(\tilde{\mathbf{x}};\phi)]
\end{align}
where ${\hat{\mathbf{x}}} := \epsilon \mathbf{x} + (1-\epsilon)\tilde{\mathbf{x}}$, for $\epsilon \sim U[0,1]$, $\mathbf{x} \sim p_r$ and $\tilde{\mathbf{x}} \sim p_\theta$.
\glnote{Mention that we use WGAN-GP in this note, but that the VO procedure can be coupled other vairants of GANs.}

@@

\glnote{add algorithm}

@@

\glnote{Implicit generative models.}
\glnote{gan + classifier ABC.}

\glnote{likelihood free inference, density ratio by classification}
\glnote{carl~\citep{cranmer2015approximating}.}
